---
title: "Simple example: compose dwellings and households"
author: "Samuel Thiriot"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8, 
  fig.height=3
)
```

Simple case to discover the principles of Direct Probabilistic Pairing. 
It illustrates how to use this method to generate populations of dwellings and households, 
both being composed inside each other depending to their characteristics.

## Load the package

```{r, fig.show='hold'}
# load the package from current directory (development mode)
library(devtools)
devtools::load_all() 

# install on a local computer: 
# library(gosp.dpp) 
```

## Input data

The data to run this case is included in the package. Let's load it and have a look to its content. 

```{r, fig.show='hold'}
data(dwellings_households)
dwellings_households
```

The data required for the generation of a synthetic population made of paired entities is made of the elements displayed there. 
We describe in detail each type of data below.

### Samples for A and B

Two samples are provided for each entities. These samples are microdata, similar to PUMS.

The sample A contains a sample of dwellings, characterised by a surface category and a cost per month:
```{r, fig.show='hold'}
dwellings_households$sample.A
head(as.data.frame(dwellings_households$sample.A))
```

The sample B contains a sample of households, characterised by a size (count of people) and an income level:
```{r, fig.show='hold'}
dwellings_households$sample.B
head(as.data.frame(dwellings_households$sample.B))
```

### Probability distributions for degrees

Input data also includes the description of how many links to create for each entity type given its characteristics. 
Here for dwellings, the count of links (in rows) depends on the attribute `surface` (in columns).
This distribution means that for dwellings having surface 1 (first column), 20% have no link (so have no link with any household,
aka empty) and 80% contain 1 household only. This table describes the fact smaller dwellings tend to host only one household, 
whilst bigger ones can contain several households.
```{r, fig.show='hold'}
dwellings_households$pdi
```

The distribution of degrees for households describes an easier situation: whatever caracteristic of the household, 
always link the household with exactly one dwelling.
```{r, fig.show='hold'}
dwellings_households$pdj
```

### Pairing probabilies

Pairing probabilities define which household and dwelling should be linked given their characteristics. 
The data proposed here describes that if we take one link randomly in the generated population, it has 20% of 
linking a small dwelling (column) with a small household (row):
```{r, fig.show='hold'}
dwellings_households$pij
```

## Solve the case

### Prepare the case

The step of preparation of data will ensure data is consistent, and is ready for resolution.

```{r, fig.show='hold'}
prepared <- matching.prepare(dwellings_households$sample.A, dwellings_households$sample.B, dwellings_households$pdi, dwellings_households$pdj, dwellings_households$pij)
```

This step will be useful and important one real cases. Il might display warnings, enrich data, or even fail if the data is not consistent.


### Resolution 

In the target population, many different variables should have consistent values. 
The resolution of the system will precisely solve the underlying system of equations.
The resolution takes as parameters relaxation parameters which help you define where you prefer the errors to take place. 

Let's first resolve the system without specific relaxation (we let the algorithm do wathever is better).
```{r, fig.show='hold'}
solved <- matching.solve(
                      prepared, nA=50000, nB=40000, nu.A=1, phi.A=1, 
                      delta.A=1, gamma=1, delta.B=1, phi.B=1, nu.B=1, verbose=FALSE)
```

Because we left the algorithm plenty of possibilities, the solver explored many possibilies.

On your actual screen, you can run `plot(solved)` and obtain a large, synthetic view of the result of the generation.
We will build here every of these plots and comment them.

The plot of errors shows a synthetic view of where the errors (more precisely, the Normalized Mean Squarred Error) 
were assigned. Here we can see the algorithm ended with errors on the count of entities $n_A$ (count of dwellings), on the pairing probabilities $p_{i,j}$, 
with no error on other characteristics.
```{r, fig.show='hold'}
plot_errors(solved)
```

We can plot the population sizes to confirm that:
```{r, fig.show='hold'}
plot_population_sizes(solved)
```

We also can investigate what happend to pairing probabilities.
```{r, fig.show='hold'}
# input probabilities
solved$inputs$pij
# solved probabilities
solved$gen$hat.pij
# display the measure of differences: positive values are probabilities in excess.
errors.pij(solved)
# we also can plot that as an heatmap. Red means "not enough created", blue "too many".
# the differences are so low they are not very visible here.
plot_errors_pij(solved) 
```


### Resolution and relaxation of parameters

Let's now imagine we would like to keep the count of entities A (dwellings), the relative frequencies for A, and the degree of connectivity. 
It means we are sure of these parameters, and accept to report errors on other characteristics. 
```{r, fig.show='hold'}
solved <- matching.solve(
                      prepared, nA=50000, nB=40000, nu.A=0, phi.A=0, 
                      delta.A=0, gamma=0, delta.B=1, phi.B=1, nu.B=1, verbose=TRUE)
```

We displayed the resolution process in order to show how it works. 
Basically, every relaxation parameter set to 0 means the initial value is taken for granted. 
Then equations are applied in order to compute values for other variables. 
If the solving leads to contradictions between values, it means the problem is overconstrained. 
Most of the time, you don't need to care about this level of detail. 


```{r, fig.show='hold'}
plot_errors(solved)
```

The plot of errors shows how the algorithm, following our request, reported the errors elsewhere (here in pairing probabilities and count of households).


Let's now try to enforce pairing probabilities by setting relaxation parameter $\gamma=0$:
```{r, fig.show='hold'}
solved <- matching.solve(
                      prepared, nA=50000, nB=40000, nu.A=1, phi.A=1, 
                      delta.A=1, gamma=0, delta.B=1, phi.B=1, nu.B=1, verbose=TRUE)
plot_errors(solved)
```

Here enforcing pairing led to the propagation of errors to many other characteristics. For instance the average degrees of dwellings was adapted.
```{r, fig.show='hold'}
plot_average_degree_A(solved)
```

We also can check what happens in detail to the distribution of degrees of probabilities, to analyze carefully which specific dwellings 
are concerned by this error. Here we observe the small dwellings have more links than planned (so there are less empy dwellings).
There are also more empty dwellings than expected.
```{r, fig.show='hold'}
# the modified distribution of degrees
solved$gen$hat.pdi
# plot it
plot_errors_pdi(solved)
```

The plot of all errors also shown how the frequencies of classes for dwellings were biased. Let's display these frequencies $f_i$, in order to analyze which are over or underrepresented.
```{r, fig.show='hold'}
plot_frequencies_A(solved)
```


## Actual generation

The actual generation actually creates the synthetic population from the computed parameters.

```{r, fig.show='hold'}
sp <- matching.generate(
                     solved, 
                     sample.A=dwellings_households$sample.A, 
                     sample.B=dwellings_households$sample.B, 
                     verbose=TRUE)
```

The corresponding result contains three main tables (R data frames) which contain dwellings, households and links between them.
```{r, fig.show='hold'}
sp
```

We can now access each data set using:
```{r, fig.show='hold'}
head(sp$pop$A)
head(sp$pop$B)
head(sp$pop$links)
```

## Assess the quality of the generation

Beyond the variables which are controlled by the algorithm, we can wonder what happened to variables which are not constrained. 
We provide a function for this plot:
```{r, fig.show='hold'}
plot_variable(sample=dwellings_households$sample.A, generated=sp$pop$A, var.name="cost.per.month")

```

## More

For more information, have a look to the conference papers related to this work, and to the other vignettes which describe the package.
